---
title: STA 380, Part 2: Exercises
author: Luke Leon
output: pdf_document
date: "2024-08-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(igraph)
library(arules)
library(arulesViz)
```

# Question 8: Association Rule Mining

I chose conservative measures for support = 0.005 and confidence = 0.1 initially so as not to limit the number of rules. After observing the full list of 1582, I narrowed down to lift > 3 and confidence > 0.1. As you can see by the resulting gephi graph, I felt these parameters result in a happy medium of not too many observations to look at while still capturing important relationships with high lift. 

Root vegetables appear to be the most common item across all baskets. A few notable items that appear in baskets together 
   - dinner essentials: chicken and vegetables
   - baking/dairy: eggs, whole milk, butter, margarine, shipped/sour cream
   - sandwich items 1: ham and white bread
   - sandwich items 2: rolls/buns, sausage, and sliced cheese


```{r}
setwd("/Users/lukeleon/Downloads")
grocery_raw <- readLines("/Users/lukeleon/Downloads/groceries.txt")

grocery_df <- data.frame(basket = grocery_raw, stringsAsFactors = FALSE)

head(grocery_df)
str(grocery_df)

# Split each grocery list into individual items
grocery_baskets <- strsplit(grocery_df$basket, ",")

# Remove duplicates 
grocery_baskets <- lapply(grocery_baskets, unique)

# Cast this variable as a special arules "transactions" class.
grocery_trans <- as(grocery_baskets, "transactions")
summary(grocery_trans)

# Now run the 'apriori' algorithm
# Look at rules with support > .005 & confidence >.1 & length (# artists) <= 4
grocery_rules <- apriori(grocery_trans, 
                         parameter = list(support = 0.005, confidence = 0.1, maxlen = 4))

# Inspect the rules
inspect(grocery_rules)

# look at subsets
inspect(subset(grocery_rules, subset = lift > 3))
inspect(subset(grocery_rules, subset = confidence > 0.6))
inspect(subset(grocery_rules, subset = lift > 3.5 & confidence > 0.4))

plot(grocery_rules)

plot(grocery_rules, measure = c("support", "lift"), shading = "confidence")

# Two-key plot
plot(grocery_rules, method = 'two-key plot')

inspect(subset(grocery_rules, support > 0.05))
inspect(subset(grocery_rules, confidence > 0.6))
inspect(subset(grocery_rules, lift > 3))

# Graph-based visualization
grocery_graph <- associations2igraph(subset(grocery_rules, lift > 3), associationsAsNodes = FALSE)
igraph::write_graph(grocery_graph, file = 'grocery_rules.graphml', format = "graphml")

```



